---
title: "p8105_HW1_nck2132"
author: "Nina Knitowski"
output: github_document
date: "09-25-2026"
---

#Question 0: Load Library
```{r}
library(tidyverse)
```

#Question 1 Load Dataset
```{r}
data("penguins", package = "palmerpenguins")
```

#1.1 Names of Dataset/Values of Important Variables
```{r}
names(penguins)
janitor::clean_names(penguins)
```
#columns and rows of dataset
```{r}
nrow(penguins)
ncol(penguins)
```
#The mean flipper length included in the skim function AND important values in the dataset
```{r}
skimr::skim(penguins)
```

The data set names are species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, year. Species, island, and sex are categorical variables and bill length, bill depth, flipper length, body mass, and year continuous variables. There are `r nrow(penguins)` rows in the dataset and `r ncol(penguins)` columns. There are `r nrow(penguins)` observations and `r ncol(penguins)` variables.
Here are some important values of the continuous variables of the table. The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` with a standard deviation of `r sd(pull(penguins, flipper_length_mm), na.rm = TRUE)`. The mean bill depth is `r mean(pull(penguins, bill_depth_mm), na.rm = TRUE)` with a standard deviation of `r sd(pull(penguins, bill_depth_mm), na.rm = TRUE)`. The mean body mass is `r mean(pull(penguins, body_mass_g), na.rm = TRUE)` with a standard deviation of `r sd(pull(penguins, body_mass_g), na.rm = TRUE)`. The mean year is `r mean(pull(penguins, year), na.rm = TRUE)` and with a standard deviation of `r sd(pull(penguins, year), na.rm = TRUE)`. The mean bill length is `r mean(pull(penguins, bill_length_mm), na.rm = TRUE)` with a standard deviation of `r sd(pull(penguins, bill_length_mm), na.rm = TRUE)`. The skimr package shows me the missing values in the dataset and does not include NA when preforming calculations of the mean and standard deviation, but for the inline code when I compute mean and SD separately I must purposefully exclude these values.


#Scatterplot of flipper_length_mm (y) vs bill_length_mm
```{r scatterplot flipper legnth vs. bill}
ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species))+geom_point()
ggsave("scatter_plot.pdf")
```
#Question 2 
#2.1 Create Data Frame
```{r Problem_2_Df}
HWProb2_df = tibble(
  norm_samp = rnorm(10, mean=1),
  norm_samp_pos = (norm_samp > 0),
  vec_char = c("My", "name", "is", "Nina", "Knitowski","and", "I", "currently","live", "in NYC"),
  vec_factor = factor(c("a", "b", "c", "c", "b", "a", "a", "b", "c", "a"))
)
```
#2.2 Try Means Using Pull
```{r try_means, warning=FALSE}
mean(pull(HWProb2_df, norm_samp))
mean(pull(HWProb2_df, norm_samp_pos))
mean(pull(HWProb2_df, vec_char))
mean(pull(HWProb2_df, vec_factor))
```
# I was able to calculate the mean of numeric vector (norm_samp), the logical vector (norm_samp_pos), but not able to calculate the mean of the character factor (vec_char) and factor vector (vec_fac).

#2.3 Convert logical, character, and factor variables to numeric
```{r convert vars, results = 'hide'}
as.numeric(pull(HWProb2_df, norm_samp_pos))
as.numeric(pull(HWProb2_df, vec_char))
as.numeric(pull(HWProb2_df, vec_factor))
```
# Logical and factor variables can be converted into numeric values, but character variables cannot. R tells us "## Warning: NAs introduced by coercion" when we try to convert a character variable into a numeric one because it is not able to convert letters and words into numbers.This helps explain why we could only take the mean of logical and numeric variables because these variables have number values. The factor vector is a categorical variable but can be converted to a numeric values for each category (1,2,3) R does not take the mean because they are numerical values used to represent the categories are not quantitative measurements. The character vector cannot be converted because it is just a set of words there is no way to turn text into a numeric value and therefore you cannot take the mean without numbers. Essentially, R does not want to take the average of character variables and categorical variables because it does not make sense. 