---
title: "p8105_HW1_nck2132"
author: "Nina Knitowski"
output: github_document
date: "09-25-2026"
---

#Question 0: Load Library
```{r}
library(tidyverse)
```

#Question 1 Load Dataset
```{r}
data("penguins", package = "palmerpenguins")
```

#1.1 Names of Dataset/Values of Important Variables
```{r}
names(penguins)
janitor::clean_names(penguins)
```
#columns and rows of dataset
```{r}
nrow(penguins)
ncol(penguins)
```
#The mean flipper length included in the skim function AND important values in the dataset
```{r}
skimr::skim(penguins)
```

#The data set names are species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, year. Species, island, and sex are categorical variables and bill length, bill depth, flipper length, body mass, and year continuous variables. There are 344 rows in the dataset and 8 columns. 
#Here are some important values of the continuous variables of the table. The mean flipper length is 200.92mm with a standard deviation of 14.06mm. The mean bill depth 43.92mm with a standard deviation is 1.97mm. The mean body mass is 4201.75 grams with a standard deviation 801.95 grams. The mean year is 2008.03 years and with a standard deviation 0.82 years. The mean bill length is 42.92mm with a standard deviation 5.46.


#Scatterplot of flipper_length_mm (y) vs bill_length_mm
```{r scatterplot flipper legnth vs. bill}
ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species))+geom_point()
ggsave("scatter_plot.pdf")
```



